---
title: "Getting started with magicaxis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with magicaxis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

The `magicaxis` package extends base R graphics with prettier scientific
axes, better defaults for log scaling, and a set of high-level helpers for
common plotting tasks. The README focuses on `magplot()`, `maghist()`, and
`magbin()`. This vignette adds examples for several other package features
that are easy to miss when browsing the reference pages.

## Core plotting workflow

`magplot()` is the usual entry point for scatter plots and line plots, while
`magaxis()` can be used directly when you need to customise axes on an
existing plot.

```{r}
library(magicaxis)

x = 10^(seq(-1, 3, length.out = 200))
y = sqrt(x) + rnorm(length(x), sd = 2)

magplot(x, y, log = "x", grid = TRUE, xlab = "Radius", ylab = "Signal")
```

If you already have a base plot, call `magaxis()` afterwards to redraw the
axes with the same house style:

```{r}
plot(x, y, axes = FALSE, xlab = "", ylab = "", log = "x")
magaxis(side = 1:2, unlog = "x", grid = TRUE, xlab = "Radius", ylab = "Signal")
box()
```

## Running summaries for noisy trends

`magrun()` is useful for summarising a noisy cloud of points before overlaying
medians, means, or quantile ranges. The output works naturally with
`magplot()` and `magerr()`.

```{r}
set.seed(1)
xy = cbind(seq(0, 4, length.out = 2000), sin(seq(0, 4, length.out = 2000) * pi))
xy[, 2] = xy[, 2] + rnorm(nrow(xy), sd = 0.25)

run = magrun(xy, bins = 12, ranges = pnorm(c(-1, 1)))

magplot(xy, col = rgb(0, 0, 0, 0.2), pch = 16, xlab = "x", ylab = "y")
lines(run$x, run$y, col = "red", lwd = 2)
magerr(run$x, run$y,
  ylo = run$y - run$yquan[, 1],
  yhi = run$yquan[, 2] - run$y,
  col = "red", length = 0
)
```

This is especially useful when you want to compare a fitted relation against
the scatter in the data, or when you want equal-count bins via `equalN=TRUE`.

## Image display and RGB composites

`magimage()` adds `magmap()`-based scaling to image plots, while
`magimageRGB()` combines separate channels into a colour composite. Both
functions accept matrices directly, but they can also work with lists that
contain `x`, `y`, and image components.

```{r}
img = outer(seq(-2, 2, length.out = 100), seq(-2, 2, length.out = 80), function(x, y) {
  exp(-(x^2 + y^2)) + 0.2 * x
})

magimage(img, stretch = "asinh", axes = TRUE)
```

```{r}
red = img
green = img^0.8
blue = img^0.6

magimageRGB(R = red, G = green, B = blue, stretch = "sqrt", saturation = 1.2)
```

Useful options to explore are:

- `stretch` and `stretchscale` for controlling contrast
- `locut` and `hicut` for clipping the dynamic range
- `qdiff=TRUE` in `magimage()` for signed difference images
- `interact=TRUE` to switch to a `plotly` viewer when the optional dependency
  is installed

## Map and sky projections

`magproj()` handles longitude/latitude plotting for astronomy and geography.
It can create the projected panel, draw grids and labels, and then accept
additional overlays with `add=TRUE`. Helpers such as `magband()` make it easy
to add wide structures like survey regions or great-circle bands.

```{r}
magproj(c(120, 240), c(-20, 20),
  type = "b", projection = "aitoff", centre = c(180, 0),
  fliplong = TRUE, labloc = c(90, -45), labeltype = "sex"
)

magband(crosseq = 180, peaklat = 0, width = 15,
  col = rgb(1, 0, 0, 0.2), border = NA
)
magproj(c(150, 180, 210), c(-10, 5, -10), type = "l", add = TRUE, col = "blue")
```

This workflow is handy when you need a publication-quality sky footprint plot
without dropping down to lower-level map projection code.

## Other helpers worth knowing about

Several functions are designed to combine naturally:

- `magcon()` for density contours on scatter plots
- `magbar()` for colour bars that match `magmap()` scaling
- `magcutout()` for extracting image subsets
- `magtri()` for triangle-style summaries of MCMC chains

The reference pages contain full argument details, but the examples above
cover the main patterns for combining the plotting helpers in practice.
