---
title: "magicaxis: Pretty Scientific Plotting"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{magicaxis: Pretty Scientific Plotting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 6
)
set.seed(42)
library(magicaxis)
```

## Overview

**magicaxis** provides functions for making pretty, publication-quality
scientific plots with base R graphics. It handles minor tick marks, log-scale
axes, 2-D density contours, image display, hexagonal binning, and triangle
(corner) plots for MCMC chains.

Since version 2.6.0, all internal argument dispatch uses
[ParmOff](https://github.com/asgr/ParmOff) instead of `do.call`.  This lets
extra `...` arguments be forwarded flexibly to each sub-function: unrecognised
arguments are silently dropped, so callers no longer need to know exactly which
parameters each inner function accepts.

---

## magaxis – pretty axis labelling

`magaxis()` adds nicely formatted tick marks and labels to an existing plot.
It supports log axes, minor ticks, Hershey fonts, and grid lines.

```{r magaxis-basic}
plot(10^(1:9), 1:9, log = "x", axes = FALSE, xlab = "", ylab = "")
magaxis(side = 1:2, xlab = "x (log scale)", ylab = "y")
```

Styling arguments such as `cex.axis` and `col.axis` are forwarded via ParmOff
to the underlying `axis()` call, while `cex.lab` and `col.lab` go to `mtext()`:

```{r magaxis-styled}
plot(1:10, (1:10)^2, axes = FALSE, xlab = "", ylab = "")
magaxis(side = 1:2,
        xlab = "x", ylab = expression(x^2),
        cex.axis = 0.8, col.axis = "steelblue",
        cex.lab  = 1.1, col.lab  = "darkred")
```

Grid lines can be toggled per axis:

```{r magaxis-grid}
plot(1:20, rnorm(20), axes = FALSE, xlab = "", ylab = "")
magaxis(side = 1:2, grid = TRUE, grid.col = "grey80", grid.lty = 2,
        xlab = "Index", ylab = "Value")
```

---

## magplot – high-level scatter plots

`magplot()` wraps `plot()` to produce pretty axes automatically.  For a simple
scatter the axis labelling, grid, and limits are handled internally:

```{r magplot-scatter}
x <- 10^(1:9)
y <- 1:9
magplot(log10(x), y, unlog = "x", xlab = "x", ylab = "y")
```

Log axes on both sides:

```{r magplot-loglog}
magplot(x, y, log = "xy", xlab = "x", ylab = "y")
```

Sigma-clipping the axis limits to focus on where the data actually are:

```{r magplot-clip}
temp <- cbind(rt(500, df = 1.5), rt(500, df = 1.5))
magplot(temp, xlim = 2, ylim = 2, xlab = "x", ylab = "y",
        main = "2-sigma clipped")
```

### z-coloured scatter

When a `z` vector is supplied, points are coloured by a colour map and an
optional colour bar is added.  The `magmap` and `magbar` arguments are filtered
via ParmOff, so unknown args are dropped cleanly:

```{r magplot-z}
n <- 300
x <- rnorm(n); y <- rnorm(n); z <- x^2 + y^2
magplot(x, y, z = z, xlab = "x", ylab = "y",
        zcol = hcl.colors(21, "YlOrRd"), dobar = TRUE)
```

---

## maghist – histogram with statistics

`maghist()` prints a summary and draws a styled histogram.  It integrates
naturally with `magplot()` because it returns a `histogram` object that
`magplot()` detects:

```{r maghist-basic}
maghist(rnorm(1000), xlab = "Value", verbose = FALSE)
```

Log-y histogram:

```{r maghist-logy}
maghist(rnorm(1e4), log = "y", grid = TRUE, verbose = FALSE,
        xlab = "Value", ylab = "Count (log)")
```

Pass the returned object back through `magplot()`:

```{r maghist-replot}
h <- maghist(10^runif(500, 0, 3), log = "x", verbose = FALSE)
magplot(h, log = "y", xlab = "x (log)", ylab = "Count (log)")
```

---

## magcon – 2-D density contours

`magcon()` computes a 2-D kernel density estimate and draws contours at chosen
probability levels (default 50%, 68%, 95%) together with an optional image
and colour bar.

```{r magcon-basic}
x <- rnorm(500); y <- x + rnorm(500, sd = 0.5)
magcon(x, y, h = c(0.2, 0.2), xlab = "x", ylab = "y", barposition = 'topleft',
       bartitleshift = 0.5)
```

Contours only (no background image):

```{r magcon-noim}
magcon(x, y, h = c(0.2, 0.2), doim = FALSE, dobar = FALSE,
       xlab = "x", ylab = "y")
```

---

## magbin – 2-D binning

`magbin()` bins 2-D data into hexagons, squares, or triangles and colours them
by count or a user-supplied statistic:

```{r magbin-hex}
xy <- cbind(rnorm(2000), rnorm(2000))
magbin(xy, shape = "hexagon", xlab = "x", ylab = "y")
```

Square bins with a z-statistic:

```{r magbin-square-z}
magbin(xy, shape = "square",
       z = xy[, 1]^2 - xy[, 2]^2,
       colref = "zstat", sizeref = "count",
       xlab = "x", ylab = "y")
```

Log-log hexagonal bins:

```{r magbin-loglog}
xylog <- cbind(10^rnorm(2000), 10^rnorm(2000))
magbin(xylog, log = "xy", xlab = "x (log)", ylab = "y (log)")
```

---

## magimage – image display

`magimage()` displays a matrix as an image with a colour map applied via
`magmap()` and optional axes via `magaxis()`.  The internal `image()` call is
dispatched through ParmOff:

```{r magimage-basic}
z <- outer(seq(-3, 3, len = 50), seq(-3, 3, len = 50),
           function(x, y) exp(-(x^2 + y^2)))
magimage(z, xlab = "x", ylab = "y")
```

Asinh stretch to reveal faint structure (the default `stretch = "asinh"` is
well suited to images with a wide dynamic range):

```{r magimage-stretch}
z2 <- outer(seq(-3, 3, len = 50), seq(-3, 3, len = 50),
            function(x, y) exp(-x^2) + 0.05 * exp(-y^2 / 0.1))
magimage(z2, stretch = "asinh", xlab = "x", ylab = "y")
```

---

## magtri – triangle (corner) plots for MCMC chains

`magtri()` makes a triangle plot summarising posterior samples.  The diagonal
shows marginal densities, the lower triangle shows 2-D density contours
(`magcon`), and the upper triangle shows scatter plots.

### Basic triangle plot

```{r magtri-basic}
chains <- data.frame(
  alpha = rnorm(500, mean =  1, sd = 0.3),
  beta  = rnorm(500, mean = -0.5, sd = 0.5),
  gamma = rnorm(500, mean =  2,   sd = 0.8)
)
magtri(chains, samples = 300)
```

### Passing extra arguments via ParmOff

Extra `...` arguments are now forwarded by ParmOff to `magaxis`, `magcon`,
**and** `points`, so you can style each sub-function from the top-level call.
Unrecognised arguments in one sub-function are silently dropped and not
recycled, so there is no risk of "unused argument" errors:

```{r magtri-styled}
# cex.axis → magaxis (tick-label size)
# lty      → magcon  (overrides contour line types)
# col      → points  (overrides the red mean-marker colour)
magtri(chains,
       samples  = 300,
       cex.axis = 0.7,
       lty      = c(2, 1, 3),
       col      = "purple")
```

### Reference values

Supply `refvals` to mark known true (or reference) parameter values with a
blue vertical line:

```{r magtri-refvals}
magtri(chains,
       samples = 300,
       refvals = c(1, -0.5, 2))
```

### Few labels per sub-panel

Supply `majorn` to reduce the number of labelled ticks.

```{r magtri-majorn}
magtri(chains,
       samples = 300,
       refvals = c(1, -0.5, 2),
       majorn = 2)
```

### Custom axis labels

Use the `lab` argument to supply expression labels for each parameter:

```{r magtri-labs}
magtri(chains,
       samples = 300,
       lab = list(
         expression(alpha),
         expression(beta),
         expression(gamma)
       ))
```
