---
title: "magicaxis: Pretty Scientific Plotting"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{magicaxis: Pretty Scientific Plotting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 6
)
set.seed(42)
library(magicaxis)
```

## Overview

**magicaxis** provides functions for making pretty, publication-quality
scientific plots with base R graphics. It handles minor tick marks, log-scale
axes, 2-D density contours, image display, hexagonal binning, and triangle
(corner) plots for MCMC chains.

Since version 2.6.0, all internal argument dispatch uses
[ParmOff](https://github.com/asgr/ParmOff) instead of `do.call`.  This lets
extra `...` arguments be forwarded flexibly to each sub-function: unrecognised
arguments are silently dropped, so callers no longer need to know exactly which
parameters each inner function accepts.

---

## magaxis – pretty axis labelling

`magaxis()` adds nicely formatted tick marks and labels to an existing plot.
It supports log axes, minor ticks, Hershey fonts, and grid lines.

```{r magaxis-basic}
plot(10^(1:9), 1:9, log = "x", axes = FALSE, xlab = "", ylab = "")
magaxis(side = 1:2, xlab = "x (log scale)", ylab = "y")
```

Styling arguments such as `cex.axis` and `col.axis` are forwarded via ParmOff
to the underlying `axis()` call, while `cex.lab` and `col.lab` go to `mtext()`:

```{r magaxis-styled}
plot(1:10, (1:10)^2, axes = FALSE, xlab = "", ylab = "")
magaxis(side = 1:2,
        xlab = "x", ylab = expression(x^2),
        cex.axis = 0.8, col.axis = "steelblue",
        cex.lab  = 1.1, col.lab  = "darkred")
```

Grid lines can be toggled per axis:

```{r magaxis-grid}
plot(1:20, rnorm(20), axes = FALSE, xlab = "", ylab = "")
magaxis(side = 1:2, grid = TRUE, grid.col = "grey80", grid.lty = 2,
        xlab = "Index", ylab = "Value")
```

---

## magplot – high-level scatter plots

`magplot()` wraps `plot()` to produce pretty axes automatically.  For a simple
scatter the axis labelling, grid, and limits are handled internally:

```{r magplot-scatter}
x <- 10^(1:9)
y <- 1:9
magplot(log10(x), y, unlog = "x", xlab = "x", ylab = "y")
```

Log axes on both sides:

```{r magplot-loglog}
magplot(x, y, log = "xy", xlab = "x", ylab = "y")
```

Sigma-clipping the axis limits to focus on where the data actually are:

```{r magplot-clip}
temp <- cbind(rt(500, df = 1.5), rt(500, df = 1.5))
magplot(temp, xlim = 2, ylim = 2, xlab = "x", ylab = "y",
        main = "2-sigma clipped")
```

### z-coloured scatter

When a `z` vector is supplied, points are coloured by a colour map and an
optional colour bar is added.  The `magmap` and `magbar` arguments are filtered
via ParmOff, so unknown args are dropped cleanly:

```{r magplot-z}
n <- 300
x <- rnorm(n); y <- rnorm(n); z <- x^2 + y^2
magplot(x, y, z = z, xlab = "x", ylab = "y",
        zcol = hcl.colors(21, "YlOrRd"), dobar = TRUE)
```

---

## maghist – histogram with statistics

`maghist()` prints a summary and draws a styled histogram.  It integrates
naturally with `magplot()` because it returns a `histogram` object that
`magplot()` detects:

```{r maghist-basic}
maghist(rnorm(1000), xlab = "Value", verbose = FALSE)
```

Log-y histogram:

```{r maghist-logy}
maghist(rnorm(1e4), log = "y", grid = TRUE, verbose = FALSE,
        xlab = "Value", ylab = "Count (log)")
```

Pass the returned object back through `magplot()`:

```{r maghist-replot}
h <- maghist(10^runif(500, 0, 3), log = "x", verbose = FALSE)
magplot(h, log = "y", xlab = "x (log)", ylab = "Count (log)")
```

---

## magclip – automatic sigma-clipping

`magclip()` iteratively removes outliers that are unlikely given a Normal
distribution.  It is used internally by `magplot()`, `maghist()`, and `magbin()`
when a scalar `xlim`/`ylim` is supplied, but it is also useful on its own:

```{r magclip-basic}
set.seed(42)
x <- c(rnorm(500), runif(30, 5, 10))   # 500 Normal + 30 outliers
result <- magclip(x, sigma = 3)
cat("Before clipping:", length(x), "points\n")
cat("After clipping:", length(result$x), "points\n")
cat("Clipped range:", round(result$range, 2), "\n")
```

The `estimate` argument controls which tail is used to estimate the Normal
dispersion.  Use `'lo'` when contamination is on the high side only:

```{r magclip-lo}
set.seed(43)
x_hi <- c(rnorm(500), runif(40, 4, 8))
r_both <- magclip(x_hi)
r_lo   <- magclip(x_hi, estimate = 'lo')
cat("Both sides kept:", length(r_both$x), "| Low-side only kept:", length(r_lo$x), "\n")
```

---

## magrun – running statistics

`magrun()` computes running medians (or means/modes) and scatter measures along
a binned axis, ideal for overlaying a trend on noisy scatter data:

```{r magrun-basic}
set.seed(44)
n  <- 1000
xx <- seq(0, 2, len = n)
yy <- xx + rnorm(n)
magplot(xx, yy, col = "lightgrey", pch = ".", xlab = "x", ylab = "y")
run <- magrun(cbind(xx, yy), bins = 10)
lines(run$x, run$y, col = "red", lwd = 2)
lines(run$x, run$yquan[, 1], lty = 2, col = "red")
lines(run$x, run$yquan[, 2], lty = 2, col = "red")
```

The `Nscale = TRUE` flag divides the scatter by √N, giving error-in-the-mean
intervals (useful for assessing the significance of a trend):

```{r magrun-Nscale}
set.seed(45)
n  <- 500
xx <- seq(0, 1, len = n)
yy <- xx + rnorm(n)
run_sd  <- magrun(cbind(xx, yy), bins = 8, Nscale = FALSE, diff = TRUE)
run_sem <- magrun(cbind(xx, yy), bins = 8, Nscale = TRUE,  diff = TRUE)
magplot(xx, yy, col = "lightgrey", pch = ".", xlab = "x", ylab = "y")
magerr(run_sd$x,  run_sd$y,  ylo = run_sd$yquan[, 1],  yhi = run_sd$yquan[, 2],
       col = "royalblue", length = 0, lty = 2)
magerr(run_sem$x, run_sem$y, ylo = run_sem$yquan[, 1], yhi = run_sem$yquan[, 2],
       col = "red")
legend("topleft", legend = c("scatter (1σ)", "SEM (1σ)"),
       col = c("royalblue", "red"), lty = c(2, 1), bty = "n")
```

---

## magerr – error bars

`magerr()` adds symmetric or asymmetric error bars in x and/or y to an
existing plot.  Lower errors are supplied via `xlo`/`ylo` and upper errors
via `xhi`/`yhi` (defaulting to the lower value if omitted):

```{r magerr-basic}
set.seed(46)
n  <- 12
x  <- sort(runif(n))
y  <- 2 * x + rnorm(n, sd = 0.15)
magplot(x, y, xlab = "x", ylab = "y")
magerr(x, y, xlo = 0.04, ylo = 0.1, yhi = 0.2)
```

Error ellipses are drawn when `corxy` is supplied:

```{r magerr-ellipse}
set.seed(47)
n <- 8
x  <- runif(n, 0.1, 0.9)
y  <- runif(n, 0.1, 0.9)
sx <- runif(n, 0.03, 0.08)
sy <- runif(n, 0.05, 0.12)
rho <- runif(n, -0.7, 0.7)
magplot(x, y, xlab = "x", ylab = "y", pch = 16)
magerr(x, y, xlo = sx, ylo = sy, corxy = rho)
```

A shaded error polygon is produced with `poly = TRUE`:

```{r magerr-poly}
set.seed(48)
x  <- seq(0, 2 * pi, len = 30)
y  <- sin(x) + rnorm(30, sd = 0.1)
magplot(x, y, type = "l", xlab = "x", ylab = "sin(x)")
magerr(x, y, ylo = 0.15, yhi = 0.15, poly = TRUE, col = adjustcolor("steelblue", 0.3),
       border = NA)
```

---

## magcurve – draw a function curve

`magcurve()` evaluates and plots any R expression over a range.  It is a
drop-in replacement for `curve()` that uses `magplot()` for its base plot,
giving pretty axes automatically:

```{r magcurve-basic}
magcurve(sin(x), from = 0, to = 2 * pi, xlab = "x", ylab = "sin(x)")
magcurve(cos(x), add = TRUE, col = "steelblue")
legend("topright", legend = c("sin", "cos"), col = c("black", "steelblue"),
       lty = 1, bty = "n")
```

Log axes are supported:

```{r magcurve-log}
magcurve(x^2, from = 1, to = 100, log = "xy",
         xlab = "x (log)", ylab = expression(x^2 ~ "(log)"))
```

---

## magbar – colour bar legend

`magbar()` adds a standalone colour bar to any existing plot.  The `position`
argument accepts compass-style strings (`'topright'`, `'bottomleft'`, etc.);
`orient` switches between vertical (`'v'`) and horizontal (`'h'`):

```{r magbar-basic}
n <- 300
set.seed(49)
x <- rnorm(n); y <- rnorm(n); z <- x^2 + y^2
colramp <- hcl.colors(21, "YlOrRd")
colvals <- magmap(z, range = c(1, length(colramp)))$map
magplot(x, y, col = colramp[round(colvals)], pch = 16,
        xlab = "x", ylab = "y")
magbar(position = "topright", range = range(z), col = colramp,
       orient = "v", title = expression(x^2 + y^2))
```

A horizontal bar at the bottom:

```{r magbar-horizontal}
magplot(x, y, col = colramp[round(colvals)], pch = 16,
        xlab = "x", ylab = "y")
magbar(position = "bottomleft", range = range(z), col = colramp,
       orient = "h", title = "z")
```

---

`magcon()` computes a 2-D kernel density estimate and draws contours at chosen
probability levels (default 50%, 68%, 95%) together with an optional image
and colour bar.

```{r magcon-basic}
x <- rnorm(500); y <- x + rnorm(500, sd = 0.5)
magcon(x, y, h = c(0.2, 0.2), xlab = "x", ylab = "y", barposition = 'topleft',
       bartitleshift = 0.5)
```

Contours only (no background image):

```{r magcon-noim}
magcon(x, y, h = c(0.2, 0.2), doim = FALSE, dobar = FALSE,
       xlab = "x", ylab = "y")
```

---

## magbin – 2-D binning

`magbin()` bins 2-D data into hexagons, squares, or triangles and colours them
by count or a user-supplied statistic:

```{r magbin-hex}
xy <- cbind(rnorm(2000), rnorm(2000))
magbin(xy, shape = "hexagon", xlab = "x", ylab = "y")
```

Square bins with a z-statistic:

```{r magbin-square-z}
magbin(xy, shape = "square",
       z = xy[, 1]^2 - xy[, 2]^2,
       colref = "zstat", sizeref = "count",
       xlab = "x", ylab = "y")
```

Log-log hexagonal bins:

```{r magbin-loglog}
xylog <- cbind(10^rnorm(2000), 10^rnorm(2000))
magbin(xylog, log = "xy", xlab = "x (log)", ylab = "y (log)")
```

---

## magimage – image display

`magimage()` displays a matrix as an image with a colour map applied via
`magmap()` and optional axes via `magaxis()`.  The internal `image()` call is
dispatched through ParmOff:

```{r magimage-basic}
z <- outer(seq(-3, 3, len = 50), seq(-3, 3, len = 50),
           function(x, y) exp(-(x^2 + y^2)))
magimage(z, xlab = "x", ylab = "y")
```

Asinh stretch to reveal faint structure (the default `stretch = "asinh"` is
well suited to images with a wide dynamic range):

```{r magimage-stretch}
z2 <- outer(seq(-3, 3, len = 50), seq(-3, 3, len = 50),
            function(x, y) exp(-x^2) + 0.05 * exp(-y^2 / 0.1))
magimage(z2, stretch = "asinh", xlab = "x", ylab = "y")
```

---

## magtri – triangle (corner) plots for MCMC chains

`magtri()` makes a triangle plot summarising posterior samples.  The diagonal
shows marginal densities, the lower triangle shows 2-D density contours
(`magcon`), and the upper triangle shows scatter plots.

### Basic triangle plot

```{r magtri-basic}
chains <- data.frame(
  alpha = rnorm(500, mean =  1, sd = 0.3),
  beta  = rnorm(500, mean = -0.5, sd = 0.5),
  gamma = rnorm(500, mean =  2,   sd = 0.8)
)
magtri(chains, samples = 300)
```

### Passing extra arguments via ParmOff

Extra `...` arguments are now forwarded by ParmOff to `magaxis`, `magcon`,
**and** `points`, so you can style each sub-function from the top-level call.
Unrecognised arguments in one sub-function are silently dropped and not
recycled, so there is no risk of "unused argument" errors:

```{r magtri-styled}
# cex.axis → magaxis (tick-label size)
# lty      → magcon  (overrides contour line types)
# col      → points  (overrides the red mean-marker colour)
magtri(chains,
       samples  = 300,
       cex.axis = 0.7,
       lty      = c(2, 1, 3),
       col      = "purple")
```

### Reference values

Supply `refvals` to mark known true (or reference) parameter values with a
blue vertical line:

```{r magtri-refvals}
magtri(chains,
       samples = 300,
       refvals = c(1, -0.5, 2))
```

### Few labels per sub-panel

Supply `majorn` to reduce the number of labelled ticks.

```{r magtri-majorn}
magtri(chains,
       samples = 300,
       refvals = c(1, -0.5, 2),
       majorn = 2)
```

### Custom axis labels

Use the `lab` argument to supply expression labels for each parameter:

```{r magtri-labs}
magtri(chains,
       samples = 300,
       lab = list(
         expression(alpha),
         expression(beta),
         expression(gamma)
       ))
```
